# Time Series Analysis

```{r}
library('urca')
library(xts)
library('vars')
library('quantmod')
library('tseries')
options("getSymbols.yahoo.warning"=FALSE)
options("getSymbols.warning4.0"=FALSE)
```

## Multivariate time series analysis

### Cointegration

# generate the two time series of length 1000

```{r}
#Modify plot parameters
par(font.axis=2, font.lab=2)

set.seed(20140623)			#fix the random seed
N <- 1000					      #define length of simulation
x <- cumsum(rnorm(N))		#simulate a normal random walk
gamma <- 0.7				    #set an initial parameter value
y <- gamma * x + rnorm(N)	#simulate the cointegrating series
plot(x, col="blue", type='l', lwd=2, ylab='simulated values')	 		#plot the two series
lines(y,col="red", lwd=2)
legend(50,30,c("x","y"),lty=c(1,1), lwd=c(2.5,2.5),col=c("blue","red")) 
```



#statistical tests

#ADF test for the simulated individual time series

	#NULL: unit root exists
	#reject NULL if test-statistic < critical value
	
```{r}
summary(ur.df(x,type="none"))	
summary(ur.df(y,type="none"))	
```

#take a linear combination of the series

```{r}
z <- y - gamma * x				
plot(z,type='l')
summary(ur.df(z,type="none"))	
```

$z_t$ clearly seems to be a white noise process; the rejection of the unit root is confirmed by the results of ADF tests

Engle-Granger method of testing cointegration:
1. Run a linear regression yt on xt (a simple OLS estimation).
2. Test the residuals for the presence of a unit root.

#Estimate the cointegrating relationship

```{r}
coin <- lm(y ~ x -1)		#regression without constant
head(coin$resid)					#obtain the residuals
summary(ur.df(coin$resid))	#ADF test of residuals
```

## Vector autoregressive models

## VAR implementation example

#download the data

```{r}
getSymbols('SNP', from='2004-01-02', to='2014-03-31')		#get S&P 500 index data
getSymbols('MSFT', from='2004-01-02', to='2014-03-31')		#get Microsoft data
getSymbols('DTB3', src='FRED')								#3-month T-Bill interest rates
```


#chart downloaded data

```{r}
chartSeries(MSFT)
chartSeries(MSFT, theme=chartTheme('white'))
```


#obtain components of downloaded object

```{r}
Cl(MSFT) 	#closing prices
Op(MSFT)	#open prices
Hi(MSFT)	#daily highest price
Lo(MSFT)	#daily lowest price
ClCl(MSFT)	#close-to-close daily return
Ad(MSFT)	#daily adjusted closing price
chartSeries(ClCl(MSFT))	#a plotting example with shortcuts #***2078OT__01_04.png***
```


#indexing time series data

```{r}
DTB3.sub <- DTB3['2004-01-02/2014-03-31']
```

#Calculate returns

```{r}
SNP.ret  <- diff(log(Ad(SNP)))
MSFT.ret <- diff(log(Ad(MSFT)))
```

#replace NA values

```{r}
DTB3.sub[is.na(DTB3.sub)] <- 0
DTB3.sub <- na.omit(DTB3.sub)
```

#merge the three databases to get the same length

```{r}
dataDaily <- na.omit(merge(SNP.ret,MSFT.ret,DTB3.sub), join='inner')
```

#obtain monthly data

```{r}
SNP.M  <- to.monthly(SNP.ret)$SNP.ret.Close
MSFT.M <- to.monthly(MSFT.ret)$MSFT.ret.Close
DTB3.M <- to.monthly(DTB3.sub)$DTB3.sub.Close
```

#Fit a simple VAR model to the data

```{r}
var1 <- VAR(dataDaily, lag.max=4, ic="AIC")
VARselect(dataDaily,lag.max=4)
summary(var1)
var1
```




#obtain the results


```{r}
summary(var1)
var1
var1$varresult
var1$type
var1$p
var1$K
var1$obs
var1$totobs
var1$restrictions
var1$call
plot(var1) 		#Diagram of fit and residuals for each variables
coef(var1)		#concise summary of the estimated variables
residuals(var1)	#list of residuals (of the corresponding ~lm)
fitted(var1)	#list of fitted values
Phi(var1)		#coefficient matrices of VMA representation
plot(var1, plot.type='multiple')
```

#impulse responses

```{r}
var.irf <- irf(var1, ci=0.9)
plot(var.irf)
```

#forecast based on a VAR model

```{r}
var.pred <- predict(var1, n.ahead=10, ci=0.95)
plot(var.pred)
```

# SVAR

#if k=3, we need 3*4/2=6 additional retriction

```{r}
amat <- diag(3)
amat[2, 1] <- NA
amat[2, 3] <- NA
amat[3, 1] <- NA
svar1 <- SVAR(var1, estmethod='direct', Amat = amat)
irf.svar1 <- irf(svar1)
plot(irf.svar1)
```

#*****************************
# Cointegrated VAR and VECM
#*****************************

#Load data

```{r}
getSymbols('DTB3', src='FRED')			#3-months T-Bill secondary market rate
getSymbols('DTB6', src='FRED')			#6-months T-Bill secondary market rate
DTB3.sub = DTB3['1984-01-02/2014-03-31']
DTB6.sub = DTB6['1984-01-02/2014-03-31']
```

#plot data

```{r}
plot(DTB3.sub)
lines(DTB6.sub, col='red')
```

#Unit root tests. Null: there is unit root.
	

```{r}
summary(ur.df(na.omit(DTB3.sub),type="none"))	
summary(ur.df(na.omit(DTB6.sub),type="none"))
```

#********************************************************
# Built-in methods: Phillips?Ouliaris Cointegration Test
#********************************************************

```{r}
x1=as.numeric(na.omit(DTB3.sub))
x2=as.numeric(na.omit(DTB6.sub))
y = cbind(x1,x2)
```


#Null hypothesis: series are NOT cointegrated.

```{r}
po.coint <- po.test(y, demean = TRUE, lshort = TRUE)
```


# Built-in methods: ca.jo

#estimation using ca.jo() - Johansen approach

```{r}
y = cbind(na.omit(DTB3.sub),na.omit(DTB6.sub))
yJoTest = ca.jo(y, type = c("trace"), ecdet = c("none"), K = 2)
#yJoRegr = cajorls(dyTest, r=1)
```


# Custom method: Engle-Granger step-by-step

#Straight linear regression without constant ***on levels***

```{r}
cregr <- lm(x1 ~ x2)
```

#OBtain residuals

```{r}
r = cregr$residuals
```

#Obtain differenced series (numeric format)

```{r}
y = cbind(x1,x2)
dy = cbind(diff(x1),diff(x2))
```

#Generate lagged variables, and adjust time series lengths respectively

```{r}
dylag1 = lag(dy)
dylag2 = lag(lag(dy))
dy = dy[4:end(dy),]
dylag1 = dylag1[3:end(dylag1),]
dylag2 = dylag2[2:end(dylag2),]
r = r[4:(end(r)-1)]
```

#VECM model, two equations: regression on differenced lags and the error term

```{r}
lmVecmLag1Var1 <- lm(dy[,1] ~  dylag1[,1] + dylag1[,2] + r )
lmVecmLag1Var2 <- lm(dy[,2] ~  dylag1[,1] + dylag1[,2] + r )
```



